---
title: "Diamonds"
author:
  - name: "Quinty Boer"
    orcid: 0009-0001-7211-9372
format:
  revealjs:
    slide-number: true
    logo: "logo.png"
    scrollable: true
    multiplex: true
    embed-resources: true
    transition: fade
    progress: false
editor: source
bibliography: bibliography.bib
---

## Diamonds are shiny ...

![Diamond image from Unsplash [@Dan_2022]](daniel-dan-gUkgk1GIGdQ-unsplash.jpg){fig-align="center"}

## ... and sparkly

![Diamond gif from Tenor [@AnitaCruz2324_2025]](diamond.gif){fig-align="center"}

## Let's look at the diamonds dataset from ggplot2

```{r}
library(ggplot2)
data(diamonds)
set.seed(42)
diamonds_sample <- diamonds[sample(nrow(diamonds), 5000), c("carat", "cut", "depth", "x", "y", "z", "price")]

library(DT)
datatable(diamonds_sample, options = list(pageLength = 3), 
          class = "compact", filter = "top", width = "95%")
```

## Histogram of diamond carat

Size matters!

::: columns

::: {.column width="50%"}


```{r cache=TRUE, label = histogram_10bins}
ggplot(diamonds_sample, aes(x = carat)) +
  geom_histogram(bins = 10, fill = "blue", color = "white") +
  theme_minimal() +
  theme(text = element_text(size = 30)) +
  labs(title = "", x = "Carat", y = "Count")
```

A histogram of diamond weight (carat), using 10 bins.

:::

::: {.column width="50%"}

```{r cache=TRUE, label=histogram_100bins}
ggplot(diamonds_sample, aes(x = carat)) +
  geom_histogram(bins = 100, fill = "red", color = "white") +
  theme_minimal() +
  theme(text = element_text(size = 30)) +
  labs(title = "", x = "Carat", y = "Count")
```

A histogram of diamond weight (carat), using 100 bins.

:::

:::

## How about a linear regression?

\

The regression equation:

\

$Y_{\text{price}} = \beta_0 + \beta_1 X_{\text{carat}} + \beta_2 X_{\text{cut}} + \\ \beta_3 X_{\text{depth}} + \beta_4 X_{\text{x}} + \beta_5 X_{\text{y}} + \beta_6 X_{\text{z}}$

## R code for the regression

\

```{r eval=FALSE, echo=TRUE}
# load diamonds dataset
library(ggplot2)
data(diamonds)

# randomly sample 5000 diamonds
set.seed(42)
diamonds_sample <- diamonds[sample(nrow(diamonds), 5000),
                            c("carat", "cut", "depth", "x",
                              "y", "z", "price")]

# fit a linear regression model
model <- lm(price ~ carat + cut + depth + x + y + z,
            data = diamonds_sample)
summary(model)
```

The dataset used in this analysis is included in the `ggplot2` package [@ggplot2].

## References

::: {#refs}

:::

