---
title: "Reproducible Simulation of Boulesteix et al. (2020)"
author: "Quinty Boer"
format: pdf
bibliography: references.bib
execute: 
  echo: true
  output: false
project:
  execute-dir: project
---

This document contains the steps and R code from a statistical simulation example given by @boulesteix_introduction_2020, in a way that should be reproducible. The simulation example investigates the impact of measurement error on a continuous dependent variable (HbA1c) and confounding variable (BMI), using NHANES data. The paper and the supplemental material with the R code ccan be accessed through <https://bmjopen.bmj.com/content/10/12/e039921>. Except for minor changes to file paths and the code to saving the plot (using `ggsave` to PNG instead of `savePlot` to TIFF), the original code is unaltered.

# Load libraries and read in data

The first steps include loading the required libraries `Hmisc`, `mice`, and `tidyverse`, as well as reading in the data. Details on the data can be found at <https://wwwn.cdc.gov/nchs/nhanes/>, and the complete database can be assessed through <https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?Begi-nYear=2015>. For this simulation, the data files were downloaded directly from <https://github.com/gerkovink/markup/blob/main/documents/week-4/raw_data>.

```{r}
#| warning: false

# load libraries
library(Hmisc)
library(mice)
library(tidyverse)
```

```{r}
# read data
d1 <- sasxport.get("../data/DEMO_I.xpt")
d2 <- sasxport.get("../data/BPX_I.xpt")
d3 <- sasxport.get("../data/BMX_I.xpt")
d4 <- sasxport.get("../data/GHB_I.xpt")
d5 <- sasxport.get("../data/TCHOL_I.xpt")
d1.t <- subset(d1,select=c("seqn","riagendr","ridageyr"))
d2.t <- subset(d2,select=c("seqn","bpxsy1"))
d3.t <- subset(d3,select=c("seqn","bmxbmi"))
d4.t <- subset(d4,select=c("seqn","lbxgh"))
d5.t <- subset(d5,select=c("seqn","lbdtcsi"))
d <- merge(d1.t,d2.t)
d <- merge(d,d3.t)
d <- merge(d,d4.t)
d <- merge(d,d5.t)
```

```{r}
# rename variables:
# RIAGENDR - Gender
# RIDAGEYR - Age in years at screening
# BPXSY1 - Systolic: Blood pres (1st rdg) mm Hg
# BMXBMI - Body Mass Index (kg/m**2)
# LBDTCSI - Total Cholesterol (mmol/L)
# LBXGH - Glycohemoglobin (%)
d$age <- d$ridageyr
d$sex <- d$riagendr
d$bp <- d$bpxsy1
d$bmi <- d$bmxbmi
d$HbA1C <- d$lbxgh
d$chol <- d$lbdtcsi
d$age[d$age<18] <- NA
```

```{r}
# select complete cases:
dc <- cc(subset(d,select=c("age","sex","bmi","HbA1C","bp")))
# analysis:
summary(lm(bp ~ HbA1C + age + as.factor(sex), data=dc))
confint(lm(bp ~ HbA1C + age + as.factor(sex), data=dc))
summary(lm(bp ~ HbA1C + bmi + age + as.factor(sex), data=dc))
confint(lm(bp ~ HbA1C + bmi + age + as.factor(sex), data=dc))
```

# Run simulations

In the simulation example from @boulesteix_introduction_2020, the original observations are assumed to be measured without error. New variables are created through simulation where measurement error is added to the variables HbA1c and/or BMI, with errors drawn from a normal distribution. The amount of measurement errors differs in the simulation scenarios. Each scenario is repeated 1000 times and the results are then averaged [@boulesteix_introduction_2020].

```{r}
# simulation of measurement error:
ref <- lm(bp ~ HbA1C + bmi + age + as.factor(sex), data=dc)$coef[2]
n.sim <- 1e3
perc.me.exp <- seq(0,.5,.1)
perc.me.conf<- seq(0,.5,.1)
scenarios <- expand.grid(perc.me.exp,perc.me.conf)
var.exp <- var(dc$HbA1C)
var.conf <- var(dc$bmi)
n <- dim(dc)[1]
beta.hat <- matrix(ncol=dim(scenarios)[1], nrow=n.sim)
for (k in 1:n.sim){
  #print(k)
  set.seed(k)
  for (i in 1:dim(scenarios)[1]){
    var.me.exp <- var.exp*scenarios[i,1]/(1-scenarios[i,1])
    var.me.conf <- var.conf*scenarios[i,2]/(1-scenarios[i,2])
    dc$HbA1C.me <- dc$HbA1C + rnorm(dim(dc)[1], 0, sqrt(var.me.exp) )
    dc$bmi.me <- dc$bmi + rnorm(dim(dc)[1], 0, sqrt(var.me.conf) )
    beta.hat[k,i] <- lm(bp ~ HbA1C.me + age + bmi.me + as.factor(sex), data=dc)$coef[2]
  }}
```

# Produce plot

The code below produces the Figure 2 in the paper by @@boulesteix_introduction_2020. It shows the estimates of the correlation between HbA1c and blood pressure when controlling for BMI in the different simulation scenarios.

```{r}
#| output: true

# create figure:
tot.mat <- cbind(100*scenarios,apply(beta.hat,2,mean))
colnames(tot.mat) <- c("me.exp","me.conf","estimate")
FIGURE <- ggplot(tot.mat, aes(me.exp, me.conf)) +
  geom_tile(color="white",aes(fill = estimate)) +
  geom_text(aes(label = round(estimate, 2))) +
  scale_fill_gradient2(low="#D55E00",mid="white",high = "#56B4E9", midpoint=ref) +
  labs(x=paste("% of total variance of HbA1c due to measurement error"),
       y=paste("% of total variance of BMI due to measurement error")) +
  coord_equal()+
  scale_y_continuous(breaks=unique(tot.mat[,1]))+
  scale_x_continuous(breaks=unique(tot.mat[,1]))+
  theme(panel.background = element_rect(fill='white', colour='grey'),
        plot.title=element_text(hjust=0),
        axis.ticks=element_blank(),
        axis.title=element_text(size=10),
        axis.text=element_text(size=10),
        legend.title=element_text(size=12),
        legend.text=element_text(size=10))
FIGURE
# savePlot("../results/Figure_STRATOS.tif", type="tif")
ggsave(filename = "../results/Figure_STRATOS.png", plot = FIGURE)
```

# References

::: {#refs}
:::

